Appendix S1. Linear regression models and non-linear population dynamics

### 2024-11-25 manuscript/Appendix_S1_v02.Rmd
Worked on Appendix S1

```{r figs9, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs9}Population size (logarithm) at one time step $N_{t+1}$ as a function of log-population size in the previous time step $N_t$. Also shown are change in population size $\\Delta N$ with population size and E."}
dN <- rep(NA, (t-1))
for (i in 1:(t-1)) {
  dN[i] <- log(N[i+1])-log(N[i])
}
d.dN <- as.data.frame(dN)
d.dN$Nt <- log(dat$N[1:(t-1)])
d.dN$Nt_plus_1 <- log(dat$N[2:t])
p1 <- ggplot2::ggplot(d.dN, aes(x=Nt,y=Nt_plus_1)) + geom_point() + theme_classic()

@@

```{r figs11, fig.pos = "H", fig.show="hold", fig.cap = "\\label{fig:figs11} Observed (grey) and model-fit (blue) values for population size at time t (x-axis) and t+1 (y-axis).",results='hide',fig.keep='all'}
Gradient <- constructGradient(m.2.sample,focalVariable="E",non.focalVariables=list(Nt=list(2),Esq=list(2)),ngrid=39)
# Esq is manually constructed as gradient-produced E^2
Gradient$XDataNew$Esq <- Gradient$XDataNew$E^2
predY <- predict(m.2.sample,XData=Gradient$XDataNew,expected=TRUE)
plotGradient(m.2.sample,Gradient,pred=predY,showData=T,measure="Y",main="",xlab="E_t",ylab="predicted N_t+1")
```

@@

```{r}
# Simulate initial species population growth with a second species present
N1.0 <- 10
N2.0 <- 10
r1.0 <- 1.67
r2.0 <- 1.7
alpha.11 <- 0.00125
alpha.22 <- 0.00125
alpha.12 <- 0.008
alpha.21 <- 0.008725
# model function
disc_LV_comp <- function(r1,r2,N1.0,N2.0,alpha.11,alpha.22,alpha.12,alpha.21) {
    Nt1 <- (r1*N1.0) / (1+alpha.11*N1.0+alpha.12*N2.0)
    Nt2 <- (r2*N2.0) / (1+alpha.22*N2.0+alpha.21*N1.0)
    return(c(Nt1,Nt2))
}
# Simulation of model for t time steps
t <- 40
N <- array(NA,dim=c(t,2))
N <- as.data.frame(N)
colnames(N) <- c("N1","N2")
N$N1[1] <- N1.0
N$N2[1] <- N2.0
for (i in 2:t) {
  res <- disc_LV_comp(r1=r1.0,r2=r2.0,N1.0=N[i-1,1],N2.0=N[i-1,2],alpha.11=alpha.11,alpha.22=alpha.22,alpha.12=alpha.12,alpha.21=alpha.21)
  N$N1[i] <- res[1]
  N$N2[i] <- res[2]
}
```
```{r figs12, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs12}Population size \\textit{N} over time \\textit{t} for a discrete-time logistic growth model, with parameters $r_1$ = 1.67, $r_2$ = 1.7, $N_{i,0}$ = 10, $\\alpha_{ii}$ = 0.00125, $\\alpha_{12}$ = 0.008, and $\\alpha_{21}$ = 0.008725."}
# Plot simulation: ggplot
N$time <- 1:t
dat <- melt(N, id.vars="time")
ggplot2::ggplot(dat, aes(time, value, col=variable)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray")
```
## Linear statistical model with covariate for both species

@@

```{r figs13, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs13}Population size \\textit{N} over time \\textit{t} for a discrete-time logistic growth model, with competition."}
# Plot simulation: ggplot
dat <- as.data.frame(cbind(N$N1,N$N2))
colnames(dat) <- c("N1","N2")
dat$time <- 1:t
p1 <- ggplot2::ggplot(dat, aes(time, N1)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray")
p2 <- ggplot2::ggplot(dat, aes(time, N2)) + geom_point()
p1 + p2
```

@@

```{r figs14, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs14}Population size over time (black line) with fitted values from a first-order autoregressive model (gray line) and from a linear model (orange line).",warning=FALSE}
colors <- c("data" = "black", "AR1 fit" = "gray", "lm fit"="orange")
p1 <- ggplot(data=dat,aes(x=time)) +
  geom_point(aes(y=log(N1),color="data")) +
  geom_point(aes(y=ar3.fit.n1, color = "AR1 fit"), alpha=0.5) +
  geom_point(aes(y=lm3.fit.n1, color = "lm fit"), alpha=0.5) +
  geom_line(aes(y=log(N1), color="data"), linewidth=.4) +
  geom_line(aes(y=ar3.fit.n1, color = "AR1 fit"), alpha=0.5, linewidth=.4) +
  geom_line(aes(y=lm3.fit.n1, color = "lm fit"), alpha=0.5, linewidth=.4) +
  theme_bw() +
  xlab("time") +
  ylab("N1") +
  labs(color="Legend") +
  scale_color_manual(values = colors)
p2 <- ggplot(data=dat,aes(x=time)) +
  geom_point(aes(y=log(N2),color="data")) +
  geom_point(aes(y=ar3.fit.n2, color = "AR1 fit"), alpha=0.5) +
  geom_point(aes(y=lm3.fit.n2, color = "lm fit"), alpha=0.5) +
  geom_line(aes(y=log(N2), color="data"), linewidth=.4) +
  geom_line(aes(y=ar3.fit.n2, color = "AR1 fit"), alpha=0.5, linewidth=.4) +
  geom_line(aes(y=lm3.fit.n2, color = "lm fit"), alpha=0.5, linewidth=.4) +
  theme_bw() +
  xlab("time") +
  ylab("N2") +
  labs(color="Legend") +
  scale_color_manual(values = colors)
p1+p2
```

```{r figs15, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs15}Population size (logarithm) at one time step $N_{t+1}$ as a function of log-population size of Species 1 and Species 2 in the previous time step $N_t$. Also shown is the change in population size of Species 1 as a function of $N_t$ for both species."}
dN1 <- rep(NA, (t-1))
for (i in 1:(t-1)) {
  dN1[i] <- log(N$N1[i+1])-log(N$N1[i])
}
d.dN1 <- as.data.frame(dN1)
d.dN1$Nt <- log(dat$N1[1:(t-1)])
d.dN1$Nt_plus_1 <- log(dat$N1[2:t])
p1 <- ggplot2::ggplot(d.dN1, aes(x=Nt,y=Nt_plus_1)) + geom_point()

d.dN1 <- as.data.frame(dN1)
d.dN1$Nt <- log(dat$N1[1:(t-1)])
d.dN1$Nt_plus_1 <- log(dat$N1[2:t])
d.dN1$N2 <- log(dat$N2[2:t])
p2 <- ggplot2::ggplot(d.dN1, aes(x=N2,y=Nt_plus_1)) + geom_point()
p3 <- ggplot2::ggplot(d.dN1, aes(Nt, dN1)) + geom_point()
p4 <- ggplot2::ggplot(d.dN1, aes(N2, dN1)) + geom_point()

@@

```{r}
df <- data.frame(cbind(log(dat$N1[2:t]),log(dat$N1[1:(t-1)]),log(dat$N2[1:(t-1)])))
colnames(df) <- c("Nt1","N1","N2")
m.3.lm <- lm(Nt1 ~ I(1/N1) + I(1/N2),data=df)

cor(df$Nt1, predict(m.3.lm))
summary(m.3.lm)
xx <- seq(0,100, length=1000)
prediction<-data.frame(N2=xx,N1=100)
plot(df$N2, df$Nt1, xlab = "N2", ylab="N.t+1" , pch=16)
lines(prediction$N2, predict(m.3.lm, prediction), lty=2,col="red",lwd=3)

```

```{r figs16, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs16}Population size (logarithm) at one time step $N_{t+1}$ as a function of log-population size of the second species in the previous time step $N_t$. Fitted (line) and observed (points) values are shown."}
jtools::effect_plot(m.3.lm,pred=N2,plot.points = TRUE)
ggplot(data = df, aes(x = N2, y = Nt1)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x)
```
We can already see that the linear model does not provide a useful fit to the data points (neither does taking the inverse function help). The regression parameters can predict well the population size of the other species, but they are not informative for inference. This has been observed previously (@Certain2018), and non-linear least squares models (among others) are more often used to estimate parameters for dynamics resulting from these kinds of species interaction models (@Kloppers2013; @Muhlbauer2020; @Olivenca2021). The inability of the linear models to reproduce the interaction coefficients is more clearly demonstrated in @Certain2018. However, they do indicate that the slope of the linear model can reflect the direction of effect for the interacting species.

Instead of further exploring techniques to fit to time-series derived from non-linear competition processes, we look more closely at how HMSC manages to fit data emerging from competition dynamics and make inference about species interactions. Instead of estimating full interaction coefficients, or assuming sparse interactions, they instead assume that a reduced number of linear combinations of species abundances that are most relevant to determining future growth rates for species in the community can bypass the 'curse of dimensionality' problem. This is well presented in @Ovaskainen2017. We apply that approach here:

@@

```{r}
preds = computePredictedValues(m.3.sample)
df <- data.frame(cbind(log(dat$N1[2:t]),log(dat$N2[2:t]),log(dat$N1[1:(t-1)]),log(dat$N2[1:(t-1)])))
colnames(df) <- c("Nt1","Nt2","N1","N2")
df$n1_25 <- NA
df$n1_975 <- NA
df$n2_25 <- NA
df$n2_975 <- NA
for(i in 1:39){
  red <- preds[i,,]
  a <- apply(red,1,quantile,probs=c(.025,.975))
  df$n1_25[i] <- a[1,1]
  df$n1_975[i] <- a[2,1]
  df$n2_25[i] <- a[1,2]
  df$n2_975[i] <- a[2,2]
}
plot(df$Nt1,pch=19,col="black",ylim=c(0,5))
points(df$Nt2,pch=19,col="orange")
lines(df$n1_25,col="black")
lines(df$n1_975,col="black")
lines(df$n2_25,col="orange")
lines(df$n2_975,col="orange")
```

I try again with a next example where the only difference between species is the $\alpha_{ij}$.

```{r}
# Simulate initial species population growth with a second species present
N1.0 <- 10
N2.0 <- 10
r1.0 <- 1.7
r2.0 <- 1.7
alpha.11 <- 0.01
alpha.22 <- 0.01
alpha.12 <- 0.005
alpha.21 <- 0.01
# model function
disc_LV_comp <- function(r1,r2,N1.0,N2.0,alpha.11,alpha.22,alpha.12,alpha.21) {
    Nt1 <- (r1*N1.0) / (1+alpha.11*N1.0+alpha.12*N2.0)
    Nt2 <- (r2*N2.0) / (1+alpha.22*N2.0+alpha.21*N1.0)
    return(c(Nt1,Nt2))
}
# Simulation of model for t time steps
t <- 40
N <- array(NA,dim=c(t,2))
N <- as.data.frame(N)
colnames(N) <- c("N1","N2")
N$N1[1] <- N1.0
N$N2[1] <- N2.0
for (i in 2:t) {
  res <- disc_LV_comp(r1=r1.0,r2=r2.0,N1.0=N[i-1,1],N2.0=N[i-1,2],alpha.11=alpha.11,alpha.22=alpha.22,alpha.12=alpha.12,alpha.21=alpha.21)
  N$N1[i] <- res[1]
  N$N2[i] <- res[2]
}
```
```{r}
# Plot simulation: ggplot
N$time <- 1:t
dat <- melt(N, id.vars="time")
ggplot2::ggplot(dat, aes(time, value, col=variable)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray")
```

>>>>

```{r figs9, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs9}Population size (logarithm) at one time step $N_{t+1}$ as a function of log-population size in the previous time step $N_t$ and of environmental value $E_t$. Also shown are change in population size $\\Delta N$ with population size and E."}
dN <- rep(NA, (t-1))
for (i in 1:(t-1)) {
  dN[i] <- log(N[i+1])-log(N[i])
}
d.dN <- as.data.frame(dN)
d.dN$Nt <- log(dat$N[1:(t-1)])
d.dN$Nt_plus_1 <- log(dat$N[2:t])
p1 <- ggplot2::ggplot(d.dN, aes(x=Nt,y=Nt_plus_1)) + geom_point() + theme_classic()

@@

```{r figs11, fig.pos = "H", fig.show="hold", fig.cap = "\\label{fig:figs11} Observed (grey) and model-fit (blue) values for population size at time t+1 (y-axis) as a function of environment $E_t$ (x-axis).",results='hide',fig.keep='all'}
Gradient <- constructGradient(m.2.sample,focalVariable="E",non.focalVariables=list(Nt=list(2),Esq=list(2)),ngrid=39)
# Esq is manually constructed as gradient-produced E^2
Gradient$XDataNew$Esq <- Gradient$XDataNew$E^2
predY <- predict(m.2.sample,XData=Gradient$XDataNew,expected=TRUE)
plotGradient(m.2.sample,Gradient,pred=predY,showData=T,measure="Y",main="",xlab="E_t",ylab="predicted N_t+1")
```

@@

```{r}
# Simulate initial species population growth with a second species present
N1.0 <- 10
N2.0 <- 10
r1.0 <- 1.67
r2.0 <- 1.7
alpha.11 <- 0.00125
alpha.22 <- 0.00125
alpha.12 <- 0.008
alpha.21 <- 0.008725
# model function
disc_LV_comp <- function(r1,r2,N1.0,N2.0,alpha.11,alpha.22,alpha.12,alpha.21) {
    Nt1 <- (r1*N1.0) / (1+alpha.11*N1.0+alpha.12*N2.0)
    Nt2 <- (r2*N2.0) / (1+alpha.22*N2.0+alpha.21*N1.0)
    return(c(Nt1,Nt2))
}
# Simulation of model for t time steps
t <- 40
N <- array(NA,dim=c(t,2))
N <- as.data.frame(N)
colnames(N) <- c("N1","N2")
N$N1[1] <- N1.0
N$N2[1] <- N2.0
for (i in 2:t) {
  res <- disc_LV_comp(r1=r1.0,r2=r2.0,N1.0=N[i-1,1],N2.0=N[i-1,2],alpha.11=alpha.11,alpha.22=alpha.22,alpha.12=alpha.12,alpha.21=alpha.21)
  N$N1[i] <- res[1]
  N$N2[i] <- res[2]
}
```
```{r figs12, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs12}Population size \\textit{N} over time \\textit{t} for a discrete-time logistic growth model, with parameters $r_1$ = 1.67, $r_2$ = 1.7, $N_{i,0}$ = 10, $\\alpha_{ii}$ = 0.00125, $\\alpha_{12}$ = 0.008, and $\\alpha_{21}$ = 0.008725."}
# Plot simulation: ggplot
N$time <- 1:t
dat <- melt(N, id.vars="time")
ggplot2::ggplot(dat, aes(time, value, col=variable)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray") + theme_classic()
```
## Linear statistical model with covariate for both species

@@

```{r figs13, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs13}Population size \\textit{N} over time \\textit{t} for a discrete-time logistic growth model, with competition between 2 species ($N_1$ and $N_2$)."}
# Plot simulation: ggplot
dat <- as.data.frame(cbind(N$N1,N$N2))
colnames(dat) <- c("N1","N2")
dat$time <- 1:t
p1 <- ggplot2::ggplot(dat, aes(time, N1)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray") + theme_classic()
p2 <- ggplot2::ggplot(dat, aes(time, N2)) + geom_point() + theme_classic()
p1 + p2
```

@@

```{r figs14, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs14}Population size over time (black line) with fitted values from a first-order autoregressive model (gray line) and from a linear model (orange line).",warning=FALSE}
colors <- c("data" = "black", "AR1 fit" = "gray", "lm fit"="orange")
p1 <- ggplot(data=dat,aes(x=time)) +
  geom_point(aes(y=log(N1),color="data")) +
  geom_point(aes(y=ar3.fit.n1, color = "AR1 fit"), alpha=0.5) +
  geom_point(aes(y=lm3.fit.n1, color = "lm fit"), alpha=0.5) +
  geom_line(aes(y=log(N1), color="data"), linewidth=.4) +
  geom_line(aes(y=ar3.fit.n1, color = "AR1 fit"), alpha=0.5, linewidth=.4) +
  geom_line(aes(y=lm3.fit.n1, color = "lm fit"), alpha=0.5, linewidth=.4) +
  theme_bw() +
  xlab("time") +
  ylab("log(N1)") +
  labs(color="Legend") +
  scale_color_manual(values = colors)
p2 <- ggplot(data=dat,aes(x=time)) +
  geom_point(aes(y=log(N2),color="data")) +
  geom_point(aes(y=ar3.fit.n2, color = "AR1 fit"), alpha=0.5) +
  geom_point(aes(y=lm3.fit.n2, color = "lm fit"), alpha=0.5) +
  geom_line(aes(y=log(N2), color="data"), linewidth=.4) +
  geom_line(aes(y=ar3.fit.n2, color = "AR1 fit"), alpha=0.5, linewidth=.4) +
  geom_line(aes(y=lm3.fit.n2, color = "lm fit"), alpha=0.5, linewidth=.4) +
  theme_bw() +
  xlab("time") +
  ylab("log(N2)") +
  labs(color="Legend") +
  scale_color_manual(values = colors)
p1+p2
```

```{r figs15, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs15}Population size (logarithm) in species 1 at one time step $N_{1,t+1}$ as a function of log-population size of Species 1 and Species 2 in the previous time step $N_t$. Also shown is the change in population size of Species 1 as a function of $N_t$ for both species."}
dN1 <- rep(NA, (t-1))
for (i in 1:(t-1)) {
  dN1[i] <- log(N$N1[i+1])-log(N$N1[i])
}
d.dN1 <- as.data.frame(dN1)
d.dN1$Nt <- log(dat$N1[1:(t-1)])
d.dN1$Nt_plus_1 <- log(dat$N1[2:t])
p1 <- ggplot2::ggplot(d.dN1, aes(x=Nt,y=Nt_plus_1)) + geom_point() + xlab("log(N1)") + ylab("log(N1_t+1)") + theme_classic()

d.dN1 <- as.data.frame(dN1)
d.dN1$Nt <- log(dat$N1[1:(t-1)])
d.dN1$Nt_plus_1 <- log(dat$N1[2:t])
d.dN1$N2 <- log(dat$N2[2:t])
p2 <- ggplot2::ggplot(d.dN1, aes(x=N2,y=Nt_plus_1)) + geom_point() + theme_classic() + xlab("log(N2)") + ylab("log(N1_t+1)")
p3 <- ggplot2::ggplot(d.dN1, aes(Nt, dN1)) + geom_point() + xlab("log(N1)") + theme_classic()
p4 <- ggplot2::ggplot(d.dN1, aes(N2, dN1)) + geom_point() + theme_classic() + xlab("log(N2)")

@@

```{r figs16, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs16}Population size (logarithm) in species 1 at one time step $N_{1,t+1}$ as a function of log-population size of Species 2 in the previous time step $N_t$."}
df <- data.frame(cbind(log(dat$N1[2:t]),log(dat$N1[1:(t-1)]),log(dat$N2[1:(t-1)])))
colnames(df) <- c("Nt1","N1","N2")
m.3.lm <- lm(Nt1 ~ I(1/N1) + I(1/N2),data=df)

cor(df$Nt1, predict(m.3.lm))
summary(m.3.lm)
xx <- seq(0,100, length=1000)
prediction<-data.frame(N2=xx,N1=100)
plot(df$N2, df$Nt1, xlab = "N2", ylab="N.t+1" , pch=16)
lines(prediction$N2, predict(m.3.lm, prediction), lty=2,col="red",lwd=3)
```

```{r figs17, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs17} Linear model-predicted population size (logarithm) at one time step $N_{t+1}$ as a function of log-population size of the second species in the previous time step $N_t$. Fitted (line) and observed (points) values are shown.", warning=FALSE}
jtools::effect_plot(m.3.lm,pred=N2,plot.points = TRUE)
# ggplot(data = df, aes(x = N2, y = Nt1)) +
#   geom_point() +
#   geom_smooth(method = "lm", formula = y ~ x) + theme_classic()
```

We can already see that the linear model does not provide a useful fit to the data points (neither does taking the inverse function help). The regression parameters can predict well the population size of the other species, but they are not informative for inference. This has been observed previously (@Certain2018), and non-linear least squares models (among others) are more often used to estimate parameters for dynamics resulting from these kinds of species interaction models (@Kloppers2013; @Muhlbauer2020; @Olivenca2021). The inability of the linear models to reproduce the interaction coefficients is more clearly demonstrated in @Certain2018. However, they do indicate that the slope of the linear model can reflect the direction of effect for the interacting species.

Instead of further exploring techniques to fit to time-series derived from non-linear competition processes, we look more closely at how HMSC manages to fit data emerging from competition dynamics and make inference about species interactions. Instead of estimating full interaction coefficients, or assuming sparse interactions, they instead assume that a reduced number of linear combinations of species abundances that are most relevant to determining future growth rates for species in the community can bypass the 'curse of dimensionality' problem. This is well presented in @Ovaskainen2017. We apply that approach here.

@@

```{r figs18, fig.pos = "H",echo=FALSE, fig.show="hold", fig.cap = "\\label{fig:figs18} Observed (points) and HMSC model-predicted population size (logarithm) for Species 1 (black) and Species 2 (orange), using environment $E_t$ as a fixed effect predictor and specifying temporal random effects to capture the species interactions."}
preds = computePredictedValues(m.3.sample)
df <- data.frame(cbind(log(dat$N1[2:t]),log(dat$N2[2:t]),log(dat$N1[1:(t-1)]),log(dat$N2[1:(t-1)])))
colnames(df) <- c("Nt1","Nt2","N1","N2")
df$n1_25 <- NA
df$n1_975 <- NA
df$n2_25 <- NA
df$n2_975 <- NA
for(i in 1:39){
  red <- preds[i,,]
  a <- apply(red,1,quantile,probs=c(.025,.975))
  df$n1_25[i] <- a[1,1]
  df$n1_975[i] <- a[2,1]
  df$n2_25[i] <- a[1,2]
  df$n2_975[i] <- a[2,2]
}
plot(df$Nt1,pch=19,col="black",ylim=c(0,5))
points(df$Nt2,pch=19,col="orange")
lines(df$n1_25,col="black")
lines(df$n1_975,col="black")
lines(df$n2_25,col="orange")
lines(df$n2_975,col="orange")
```

To show that the HMSC model correctly allocates variance to the contributing components of the model (here environment, intra-specific and interspecific density-dependence, implemented as temporal random effects), we simulate a new example where the only difference between species is the $\alpha_{ij}$.

```{r}
# Simulate initial species population growth with a second species present
N1.0 <- 10
N2.0 <- 10
r1.0 <- 1.7
r2.0 <- 1.7
alpha.11 <- 0.01
alpha.22 <- 0.01
alpha.12 <- 0.005
alpha.21 <- 0.01
# model function
disc_LV_comp <- function(r1,r2,N1.0,N2.0,alpha.11,alpha.22,alpha.12,alpha.21) {
    Nt1 <- (r1*N1.0) / (1+alpha.11*N1.0+alpha.12*N2.0)
    Nt2 <- (r2*N2.0) / (1+alpha.22*N2.0+alpha.21*N1.0)
    return(c(Nt1,Nt2))
}
# Simulation of model for t time steps
t <- 40
N <- array(NA,dim=c(t,2))
N <- as.data.frame(N)
colnames(N) <- c("N1","N2")
N$N1[1] <- N1.0
N$N2[1] <- N2.0
for (i in 2:t) {
  res <- disc_LV_comp(r1=r1.0,r2=r2.0,N1.0=N[i-1,1],N2.0=N[i-1,2],alpha.11=alpha.11,alpha.22=alpha.22,alpha.12=alpha.12,alpha.21=alpha.21)
  N$N1[i] <- res[1]
  N$N2[i] <- res[2]
}
```
```{r}
# Plot simulation: ggplot
N$time <- 1:t
dat <- melt(N, id.vars="time")
ggplot2::ggplot(dat, aes(time, value, col=variable)) + geom_point() + geom_hline(yintercept = ((r1.0 - 1)/alpha.11),linetype = "dashed", color = "gray") = theme_classic()
```
